\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2018-01-22 document/XSGRB_sample.tex
Notes on individual bursts

\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)}
A separate reduction of this burst has been published in \citet{Kruhler2015},
but not otherwise. Starting 9 days after the burst trigger, this is host
observation and does not contain any afterglow continuum. In 3.6 ks integration
time, we detect a faint continuum visible from 450 nm and all the way through
2100 nm, in contrast to what is found previously. The continuum from 4500 -
600 nm is detected at very low significance. If the drop at 450 nm is the
Lyman limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim
3.9$. The absence of nebular lines if due to \oii~falling in a telluric
absorption band and the rest being shifted out of the wavelength coverage.

\subsection{GRB~120815A* (z = 2.358)} 
Not a part of the statistical sample, thisburst also formed the basis of  GCN
\#13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
published in \citet{Kruhler2013}. Observations started 1.69 hours after the BAT
trigger and consist of 2.4 ks integration. A bright afterglow continuum is
detected across the entire spectral coverage of X-shooter, with a multitude of
absorption lines superposed. Absorption features from the host at $z = 2.358$
include a DLA as well as metal absorption lines from \nv, \sii, \SIii, \oi,
\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
fine structure lines from \NIii and \feii are exited local to the GRB.
Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.

@@

This spectrum has already been used in GCN
\#15250\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15250.gcn3}} and is
additionally published in \citet{Schady2015} and \citet{Kruhler2015}.
Observations of this burst began with X-shooter 3.5 hours after the
\swift~trigger. 6 ks spectroscopic integration captures a heavily dust obscured
afterglow (A$_\mathrm{V} = 5.9 \pm 0.7$\citep{Greiner2014}), with the spectrum
primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, which we as the
redshift of the GRB. This spectrum is taken under ESO programme ID:
091.A-0877(A) (PI: Schady).

@@

%GCN Circular #16194 
%T. Kruehler (ESO Santiago), D. Malesani (DARK/NBI), A. de Ugarte Postigo (IAA/CSIC and DARK/NBI), A. Melandri (INAF/OABr), J. P. U. Fynbo (DARK/NBI), report on behalf of a larger collaboration: 
%We observed the optical afterglow of GRB 140430A (Siegel et al., GCN 16190) with the ESO VLT equipped with X-shooter. Observations started at 23:09 UT on 2014 April 30 (2.5 hr after the GRB), still in twilight, and two 10-minute spectra were secured. The covered wavelength range is 3000-25000 AA. 
%We detect a clear continuum, and several absorption features, which we interpret as due, among others, to Si II, C IV, Al II, Fe II, and Mg II at a common redshift z = 1.60. In the infrared arm, we also detect the [O III] 5008 nebular line in emission at the same redshift.

\subsection{GRB~140506A  (z = 0.889)}	
The data presented here has formed the basis of GCN \#
16214\footnote{\url{http://gcn.gsfc.nasa.gov/other/140506A.gcn3}} and is
published in \citet{Fynbo2014} and \citet{Heintz2017a}. The observations
consists of $4\times 600$ sec at 8.8 and 33 hours after the burst. We detect
absorption features from a wide range of ions, together with molecular
absorption from CH+, all at $z=0.889$. The optical/near-infrared afterglow
reveal an unusual steep extinction curve which is found to be caused by dust
immediate to the burst.

\subsection{GRB~140515A (z = $\sim6.32$?)}

@@

>>>>

\subsection{GRB~120805A (z $\sim$ 3.9 NEW!!!)}
A separate reduction of this burst has been published in \citet{Kruhler2015},
but is not used otherwise. Starting 9 days after the burst trigger, this is host
observation and does not contain any afterglow continuum. In 3.6 ks integration
time, we detect a faint continuum visible from 450 nm and all the way through
2100 nm, in contrast to what is found previously. The continuum from 4500 -
600 nm is detected at very low significance. If the drop at 450 nm is the
Lyman limit, this fits with Lyman alpha at $\sim$ 600 nm, giving $z \sim
3.9$. The absence of nebular lines if due to \oii~falling in a telluric
absorption band and the rest being shifted out of the wavelength coverage.

\subsection{GRB~120815A (z = 2.358)} 
Not a part of the statistical sample, this burst also formed the basis of  GCN
\#13649\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/13649.gcn3}} and is
published in \citet{Kruhler2013}. Observations started 1.69 hours after the BAT
trigger and consist of 2.4 ks integration. A bright afterglow continuum is
detected across the entire spectral coverage of X-shooter, with a multitude of
absorption lines superposed. Absorption features from the host at $z = 2.358$
include a DLA as well as metal absorption lines from \nv, \sii, \SIii, \oi,
\civ, \SIiv, \feii, \alii, \aliii, \mnii, \mgii, and \mgi. Additionally
fine structure lines from \NIii and \feii are exited local to the GRB.
Intervening systems are found at $z = 1.539$, $z = 1.693$, and $z = 2.00$.

@@

This spectrum has already been used in GCN
\#15250\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/15250.gcn3}} and is
additionally published in \citet{Schady2015} and \citet{Kruhler2015}.
Observations of this burst began with X-shooter 3.5 hours after the
\swift~trigger. 6 ks spectroscopic integration captures a heavily dust obscured
afterglow (A$_\mathrm{V} = 5.9 \pm 0.7$\citep{Greiner2014}), with the spectrum
primarily dominated by host emission lines. All the nebular lines (\oii, \hg,
\hb, \oii, \ha, \nii, \sii) are well detected at $z = 0.347$, as well as those
from \pad, \pag, and \pab. We take this as the redshift of the GRB. This
spectrum is taken under ESO programme ID: 091.A-0877(A) (PI: Schady).

@@

A separate reduction of this burst has been published in \citet{Kruhler2015},
and additionally the spectrum has also been used in GCN
\#16194\footnote{\url{http://gcn.gsfc.nasa.gov/gcn3/16194.gcn3}}. Observations
for this burst began in twilight, 2.5 hours after the BAT trigger and lasted for
$2\times600$ s. The spectrum contains a moderate S/N GRB afterglow continuum all
the way through the spectroscopic arms of X-shooter. We identify absorption
features in the afterglow continuum from \SIii, \civ, \alii, \feii, and \mgii,
and emission lines from \oii~and \oiii~- all at $z = 1.601$ which is likely the
redshift of the GRB.

\subsection{GRB~140506A  (z = 0.889)}

The data presented here has formed the basis of GCN
\#16217\footnote{\url{https://gcn.gsfc.nasa.gov/gcn3/16217.gcn3}} and is
published in \citet{Fynbo2014}, \citet{Kruhler2015}, and \citet{Heintz2017a}.
The observations consists of $4\times 600$ s at 8.8 and 33 hours after the
burst. We detect absorption features from a wide range of ions, together with
molecular absorption from CH+, all at $z=0.889$. The optical/near-infrared
afterglow reveal an unusual steep extinction curve which is found to be caused
by dust immediate to the burst.

\subsection{GRB~140515A (z = 6.327)}

This spectrum has previously been used in \citet{Melandri2015}, but is not
published elsewhere. Starting 15.5 hours after the \swift~trigger, $8\times600$
s spectroscopic integration captures this very high redshift GRB afterglow. The
Gunn-Peterson trough is visible against the GRB afterglow continuum, starting at
8900 \AA, which along with absorption lines from the \mgii~doublet securely sets
the redshift of this GRB at $z = 6.327$. From the red wing of the \lya-profile
we measure \nh~$=19.0 \pm 0.5$, which is very low compared to the measured
distribution of $N_{\mathrm{HI}}$.

@@

This spectrum forms the basis for GCN
\#16401\footnote{\url{http://gcn.gsfc.nasa.gov/other/16401.gcn3}}, but is not
published elsewhere. $4\times600$ s spectroscopic integration, starting 3.8
hours after the BAT trigger catches this GRB afterglow. The continuum is
detected at moderate S/N in both the VIS and NIR arms of X-shooter, heavily
affected by absorption features. We identify the the lines belonging to \lya,
\SIii, \cii, \cii$^*$, \alii, \aliii, \feii, and \mgii~at a consistent redshift
of $z = 4.233$. From the \lya~absorption trough we measure \nh~$=21.3 \pm 0.3$.
